---
title: "MCMC_HW1"
author: "Suman"
date: "03/10/2021"
output:
  pdf_document: default
  html_document:
    theme: readable
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


## [HW1] Generate 10,000 samples from Gamma(4,2) and Gamma(6,3) and then draw histograms.

$$
\begin{aligned}
Gamma(4,2) & = \sum_{i=1}^4 Gamma(1,2)\quad (i.i.d.) \\  
            & = \sum_{i=1}^4 Exp(2)
\end{aligned}
$$

```{r Gamma distribution1}
alpha = 4 ; beta = 2
u = matrix(runif(10000*alpha), ncol=alpha)

# Using inverse cdf method for exponential distribution with parmater 2
x = -(1/beta)*log(1-u)

# Sum of 4 exp(2)
x_sum = rowSums(x)

# Histogram
hist(x_sum, nclass=100, main="Histogram of 10,000 samples from Gamma(4,2)")

```

$$
\begin{aligned}
Gamma(6,3) &= \sum_{i=1}^6 Gamma(1,3)\quad (i.i.d.) \\  
           &= \sum_{i=1}^6 Exp(3)  
\end{aligned}
$$


```{r Gamma distribution2}
alpha=6 ; beta=3
u = matrix(runif(10000*alpha), ncol=alpha)

# Using inverse cdf method for exp(3)
x = -(1/beta)*log(1-u)

# Sum of 6 exp(3)
x_sum = rowSums(x)

# Histogram
hist(x_sum, nclass=100, main="Histogram of 10,000 samples from Gamma(6,3)")
```

First, we know that the pdf of Cauchy(0,1) is 
$$f(x) = \frac{1}{\pi(1+x^2)}\quad x\in\Re$$  
Then the cdf(cumulative distribution function) of the Cauchy(0,1) is

$$
\begin{aligned}
\int_{- \infty}^x \frac{1}{\pi(1+t^2)}dt &= \frac{1}{\pi}arctan(t)\mid_{-\infty}^x\\ & =\frac{1}{\pi}arctan(x) + \frac{1}{2}
\end{aligned}
$$

By using inverse cdf method, we can get samples from Cauchy(0,1)
$\\$

$u = \frac{1}{\pi}arctan(x) + \frac{1}{2} \\$
$\\$
$x = tan(\pi(u-\frac{1}{2})) \\$  


In this example, $\pi$ is regarded as 3.14.


```{r Cauchy distribution}
u = runif(10000)
pi = 3.14
x = tan(pi*(u-1/2))
hist(x, nclass=100, main="Histogram of 10,000 samples from Cauchy(0,1)")
```


## [HW 2] Suppose our target density is the triangle density where


$$
F(x) = 
\begin{cases} 
0 & if\ x < 0 \\
4x^2 & if\ 0 \leq x < 0.25 \\
\frac{8}{3}x - \frac{4}{3}x^2 - \frac{1}{3} & if\ 0.25\leq x\leq 1 \\
1 & if\ x > 1 
\end{cases}
$$

### (a) Generate random number using inverse-CDF

For using inverse-CDF, we need to transform F(x) to be inversed.\\
When we set $F(x) = u$ and solve for $x$, 

$$
F^{-1}(u) =
\begin{cases}
\frac{\sqrt{u}}{2} & if\ 0\leq u < 0.25 \\
1 - \frac{\sqrt{3(1-u)}}{2} & if\ 0.25\leq u \leq 1 \\
\end{cases}
$$


Even if we can not get the solutions when u is not in $[0,1]$, it is okay that u follows $Unif(0,1)$

```{r inverse-CDF of F(x)}
# sample size
n <- 10000

# inverse CDF
u <- runif(10000)
inverse.cdf <- function(u) {
    if (u >= 0 && u < 0.25)
        sqrt(u)/2
    else if (u >=0.25 && u <=1)
        1 - sqrt(3*(1-u))/2 }

# Generate random number 
x <- unlist(lapply(u, inverse.cdf))

# Show the 100 samples of generated random numbers
head(x, 100)

# Histogram 
hist(x, nclass=100, main="Histogram of 10,000 samples from F(x)")
```

### (b) Suppose g follows unif(0,1) and e = 3g. Generate samples using rejection sampling.

First, we should get the target density $f(x)$ from $F(x)$.

$$
f(x) = 
\begin{cases}
8x & if\ 0\leq x < 0.25 \\
\frac{8}{3} - \frac{8}{3}x & if\ 0.25\leq x \leq 1 \\
0 & o.w.
\end{cases}
$$
Using rejection sampling method... 

```{r rejection sampling}

## 1. Setup

# 1-0. desired sample size and sample space
desired_size <- 10000
samples <- c()

# 1-1. target density f(x)

target.x <- function(x) { 
    if ((x >=0) && (x < 0.25))
        8*x
    else if ((x >=0.25) && (x<=1))
        8/3 - 8/3*x
    else
        0
}

# 1-2. proposal density 
proposal.x <- function(x) {
    if (x>=0 && x<=1)
        1
    else 
        0
}

# 1-3. envelope density e(x) = 3g(x)
envelope.x <- function(x) {
    if (x>=0 && x<=1)
        3
    else
        0
}


## 2. Rejection sampling

# 2-1. Sample y from g(x)
# y <- runif(1)

# 2-2. Sample u from Unif(0,1)
#u <- runif(1) 

# 2-3. Calculation of rejection probability
# rej_prob <- target.x / envelope.x

### reject y if u > rej_prob

# This process should be repeated until we have samples of the desired size(10,000)
curr_size <- 0
while(curr_size < desired_size) {
    y <- runif(1)
    u <- runif(1)
    rej_prob <- target.x(y) / envelope.x(y)
    
    if (u <= rej_prob) {
        samples <- c(samples, y)
        curr_size <- curr_size + 1
    } 
}

# 3. Compare the target density and distribution of samples of rejection sampling.
value <- c()
res <- c()

for (i in sort(runif(10000))) 
{res <- c(res, target.x(i))
value <- c(value, i)}


hist(samples, nclass= 100, freq=FALSE, main="Histogram of rejection sampling samples")
lines(value ,res, type="l", col="red")
legend("topright",c("Target density"), col=("red"), lty=1)

```

We can get samples which have almost same distribution with target density by using rejection sampling method. Also, the result of rejection sampling is almost same with the result of invese-CDF method.  



## [Example] Sampling a Bayesian Posterior
```{r Bayesian Posterior}

# set values
x <- c(8,3,4,3,1,7,2,6,2,7)
n <- 10000
lambda.samp <- rep(NA, n)
xbar = mean(x)

# proposal density
iter = 1
total = 1

while(iter <= n){
  lambda = exp(rnorm(1, log(4), 0.5))
  u = runif(1,0,1)
  ratio = exp(sum(dpois(x, lambda, log=TRUE)) - sum(dpois(x, xbar, log=TRUE)))
  if (u <= ratio){
    lambda.samp[iter] = lambda
    iter = iter+1
  }
  total = total+1
}

hist(lambda.samp, nclass=100, main= "Histogram of samples")
n/total
```