---
title: "MCMC_HW2"
author: "Suman"
date: "03/17/2021"
output:
  pdf_document: default
  html_document:
    theme: readable
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
```


### [HW 1] Generate 10,000 samples from Weibull(3, 0.3) and draw the histogram.

Let $X \sim Weibull(\lambda, k)$ and $u \sim Unif(0,1)$.  
Then by using inverse-cdf method, we can get $$x = \lambda[-log(1-u)]^\frac{1}{k}$$
    
And using this fact, we can get the result below.

```{r Weibull distribution}
lambda = 3 ; k = 0.3
u = runif(10000)
x = lambda*(-log(1-u))^(1/k)
hist(x, nclass=500, xlim= c(0, 300),main="Histogram of 10,000 samples from Weibull(3, 0.3)")
```


### [HW 2] Polar Methods for Generating Normal random variables.

Let X and Y follows standard normal distribution. Let $r$ and $\theta$ denote the polar coordinate of the vector $(X,Y)$. $$r^2 = X^2 + Y^2\quad tan\theta = \frac{Y}{X}$$
    
Since X and Y are independent, the joint density of X and Y is 
$$f(x,y) = \frac{1}{2\pi}e^{-\frac{(x^2+y^2)}{2}}$$
    
Using polar method, X and Y can be expressed as $rcos\theta, rsin\theta$ respectively. 
$$x^2+y^2 = r^2cos^2\theta + r^2sin^2\theta = r^2$$
    
Therefore, $f(x,y)$ is transformed to $f(r,\theta)$
$$f(r, \theta) = \frac{1}{2\pi}\times\frac{1}{2}e^{-\frac{r^2}{2}}$$
    
And we can think that $\frac{1}{2\pi}$ as the distribution of $\theta$ and it follows $Unif(0,2\pi)$ and $r^2 \sim exp(\frac{1}{2})$
    
By inverse-CDF method, let $u \sim Unif(0,1)$. Then we can the result below.
$$\theta = 2\pi u,\quad r = \sqrt{-2logu}$$

$$
\begin{aligned}
x = \sqrt{-2logu_1}cos2\pi u_2 \\
y = \sqrt{-2logu_1}sin2\pi u_2
\end{aligned}
$$
    
```{r polar method for generating normal r.v.}
u1 = runif(10000)
u2 = runif(10000)
r = sqrt(-2*log(u1))
theta = 2*pi*u2
x = r*cos(theta)
y = r*sin(theta)

# histogram
hist(c(x,y), nclass=100, main="Sampling result using polar method")
```


### [HW 3] 



**(1) Find a density for that ARMS is applicable**  

I will use t-mixture distribution in this section.    
  
$$ x \sim 0.3t(1) + 0.7t(4)$$


**(2) Apply arms command in armspp package**

```{r arms}
library(armspp)

# mixture of t distributions
dtmixture <- function(x) {
    parts <- log(c(0.3, 0.7)) + dt(x, df=c(1,4), log=TRUE)
    log(sum(exp(parts - max(parts)))) + max(parts)
}
curve(exp(Vectorize(dtmixture)(x)), -10, 10)
```


**(3) Generate results**  
  
```{r results}
samples <- arms(1000, dtmixture, -1000, 1000)
hist(samples, freq=FALSE, nclass = 100, main="Histogram of mixture of t-distributions")
curve(exp(Vectorize(dtmixture)(x)),-10,10,col="red", add=TRUE)
```


